---
title: "Flights R&D"
author: "Bhanu Kiran"
date: "23/01/2022"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(stringr)
```

# Meta-Data

1. ECTL_ID - Unique numeric identifier for each flight in Eurocontrol PRISME DWH  
2. ADEP ICAO -  airport code for the departure airport of the flight. The ICAO airport code or
location indicator is a four-letter alphanumeric code designating each airport around
the world. These codes are defined by the International Civil Aviation Organization,
and published in ICAO Document 7910: Location Indicators.  
3.ADEP Latitude -  Latitude of departure airport in decimal degrees  
4. ADEP Longitude - Longitude of departure airport in decimal degrees  
5. ADES ICAO - airport code for the destination airport of the flight. The ICAO airport code or
location indicator is a four-letter alphanumeric code designating each airport around
the world. These codes are defined by the International Civil Aviation Organization,
and published in ICAO Document 7910: Location Indicators.
If a flight is diverted, then the ADES will be the actual airport where it landed.  
6. ADES Latitude - Latitude of destination airport in decimal degrees.  
7. ADES Latitude - Latitude of destination airport in decimal degrees  
8. Filed Arrival Time - Time of arrival (UTC) based on the last filed flight plan. It is the time at which the
aircraft will land at the aerodrome according to the planned profile calculated for the
flight.  
9. Actual Off-Block Time - Off-Block Time (UTC) based on the ATFM-updated flight plan. The time that an aircraft
departs from its parking position. This time may be known from flight data updates received by NM, or in the absence of
such updates may be calculated from the known take-off time minus a standard taxi
time value for the airport.  
10.Actual Arrival
Time - 
Time of arrival (UTC) based on the ATFM-updated flight plan. It is the time at which the
aircraft lands at the aerodrome, reflecting the best picture that NM has based on
available radar updates, ATFM messages received etc.  
11. AC Type - The ICAO aircraft type designator is a two-, three- or four-character alphanumeric code
designating every aircraft type that may appear in flight planning. These codes are
defined by the International Civil Aviation Organization, and published in ICAO
Document 8643 Aircraft Type Designators.  
12. AC Operator - Three-letter ICAO operator code. Aircraft operator codes are defined by ICAO and
published in Document 8585. If the operator is unknown, not provided in the flight
plan or not present in Document 8585 the value will be "ZZZ".  
13. AC Registration - Aircraft registration. In accordance with the Convention on International Civil Aviation,
all civil aircraft must be registered with a national aviation authority (NAA) using
procedures set by each country. Every country, even those not party to the Chicago
Convention, has an NAA whose functions include the registration of civil aircraft. An
aircraft can only be registered once, in one jurisdiction, at a time. The NAA allocates a
unique alphanumeric string to identify the aircraft, which also indicates the nationality
of the aircraft, and provides a legal document called a Certificate of Registration, one
of the documents which must be carried when the aircraft is in operation.
14. ICAO Flight Type - ICAO Flight Type: S – Scheduled, N - Non-scheduled commercial operation  
15. STATFOR Market
Segment - Market segment definitions can be found in
http://www.eurocontrol.int/sites/default/files/content/documents/officialdocuments/
facts-and-figures/statfor/statfor-market-segments-rules-for-sid-2016-
definition.xls
16. Requested FL -  Requested cruising flight level from the flight plan.
17. Actual Distance
Flown (nm) - 
Distance flown in nautical miles, corresponding to the ‘actual’ profile below.  


# TODO:

1. import all the data sets. 
2. join the datasets where possible. check python GitHub code. https://github.com/eeftychiou/FuelCost/blob/master/preprocess.py
3. according to the link create functions accordingly. EG: loading file, joining files, refer the python code.
4. The functions have math calculations so do them.



# Importing data sets
```{r}
setwd("/Users/bhanukiran/Desktop/Fiverr gig Tashayu" )
directory<-('./flights data/')
files <- list.files(directory, full.names=T)
data.list <- lapply(files, read.csv)
flights_df <- do.call(rbind, data.list)
dim(flights_df)
#flight1 <- read_csv('Flights_20190301_20190331.csv')
#flight2 <- read_csv('Flights_20190601_20190630.csv')
#flight3 <- read_csv('Flights_20190901_20190930.csv')
#flight4 <- read_csv('Flights_20191201_20191231.csv')
```


### Changing column names of the columns



```{r}
#setColNames <- function(flights_df) {
colnames(flights_df) <- c("ECTRL_ID", "ADEP", "ADEP_Latitude", "ADEP_Longitude", "ADES", "ADES_Latitude", "ADES_Longitude",
                   "FILED_OFF_BLOCK_TIME", "FILED_ARRIVAL_TIME",
                   "ACTUAL_OFF_BLOCK_TIME", "ACTUAL_ARRIVAL_TIME", "AC_Type", "AC_Operator", "AC_Registration",
                   "ICAO_Flight_Type", "STATFOR_Market_Segment", "Requested_FL", "Actual_Distance_Flown")

flights_df$STATFOR_Market_Segment <- as.factor(flights_df$STATFOR_Market_Segment)
flights_df$AC_Type <- as.factor(flights_df$AC_Type)
flights_df$ADES <- as.factor(flights_df$ADES)
flights_df$ADEP <- as.factor(flights_df$ADEP)
flights_df$AC_Operator <- as.factor(flights_df$AC_Operator)
flights_df$Actual_Distance_Flown <- as.numeric(flights_df$Actual_Distance_Flown)
#return(flights_df)
#}
```

```{r}
#flights_df<-setColNames(flights_df)
```

```{r}
colnames(flights_df)
```

```{r}
head(flights_df)
```


### Load acperf data

```{r}
#preProcessAceperf <- function(flights_df){

header_list <- c("KEY","BK_AC_TYPE_ID","ICAO_TYPE_CODE","AC_CATEGORY","EQV_TYPE","EQV_NAME","ICAO_ENGINE_DESC","CO2_COEFF","MASS","BAND_FROM_NM","BAND_TO_NM","BAND_FLOOR_NM","FUEL_TOT","FUEL_TOT_MARG_RATE","CORR_FACTOR","CALC_RETURN_CODE","TINV","S","N","X_BAR","SXX","E","ERROR_TYPE","MASS_RATIO","ERROR_RATE_FUEL_PER_NM","AO_FUEL_VERSION_ID","CREA_DATE","CREA_NOTE","VALID_FROM","VALID_TO")

directory<-('./aceperfDB data/')
files <- list.files(directory, full.names=T)
data.list <- lapply(files, read_xlsx)
aceperf_df <- do.call(rbind, data.list)
colnames(aceperf_df) <- header_list

aceperf_df$KEY <- as.factor(aceperf_df$KEY)
aceperf_df$CO2_COEFF <- as.numeric(aceperf_df$CO2_COEFF)
aceperf_df$FUEL_TOT <- as.numeric(aceperf_df$FUEL_TOT)
aceperf_df$FUEL_TOT_MARG_RATE <- as.numeric(aceperf_df$FUEL_TOT_MARG_RATE)
aceperf_df$CORR_FACTOR <- as.numeric(aceperf_df$CORR_FACTOR)

flights_df$AC_Type <- gsub(" ","",paste(flights_df$AC_Type, '----'))
colnames(aceperf_df)[1] <- 'AC_Type'

flights_df <- full_join(flights_df, aceperf_df)



flights_df$FUEL <- ((flights_df$FUEL_TOT + (flights_df$Actual_Distance_Flown * flights_df$FUEL_TOT_MARG_RATE)) * flights_df$CORR_FACTOR )
flights_df$EMISSIONS <- flights_df$CO2_COEFF * flights_df$FUEL

#return(flights_df)

#}
```

```{r}
#flights_df<-preProcessAceperf(flights_df)
```

```{r}
head(aceperf_df)
```


```{r}
head(flights_df)
```



### Create categories

```{r}

createCategories <- function(flights_df){
directory<-('./ICAOPrefix data/')
files <- list.files(directory, full.names=T)
data.list <- lapply(files, read_xlsx)
icaoPrefixCategories <- do.call(rbind, data.list)
dim(flights_df)
flights_df$ADEP_PREFIX <- substring(flights_df$ADEP, 1,2)
flights_df$ADES_PREFIX <- substring(flights_df$ADES, 1,2)

#fix flights to/from US, CANADA and Australia

flights_df$ADEP_PREFIX <- ifelse(substring(flights_df$ADEP_PREFIX,1,1) == 'K', 'K', substring(flights_df$ADEP_PREFIX,1,2))
flights_df$ADES_PREFIX <- ifelse(substring(flights_df$ADES_PREFIX,1,1) == 'K', 'K', substring(flights_df$ADES_PREFIX,1,2))

flights_df$ADEP_PREFIX <- ifelse(substring(flights_df$ADEP_PREFIX,1,1) == 'C', 'C', substring(flights_df$ADEP_PREFIX,1,2))
flights_df$ADES_PREFIX <- ifelse(substring(flights_df$ADES_PREFIX,1,1) == 'C', 'C', substring(flights_df$ADES_PREFIX,1,2))

flights_df$ADEP_PREFIX <- ifelse(substring(flights_df$ADEP_PREFIX,1,1) == 'Y', 'Y', substring(flights_df$ADEP_PREFIX,1,2))
flights_df$ADES_PREFIX <- ifelse(substring(flights_df$ADES_PREFIX,1,1) == 'Y', 'Y', substring(flights_df$ADES_PREFIX,1,2))

#fix flights to/from China
flights_df$ADEP_PREFIX <- ifelse(substring(flights_df$ADEP_PREFIX,1,1) == 'Z' & substring(flights_df$ADEP_PREFIX,1,2) != 'ZK' & substring(flights_df$ADEP_PREFIX,1,2) != 'ZM', 'Z', substring(flights_df$ADEP_PREFIX,1,2))
flights_df$ADES_PREFIX <- ifelse(substring(flights_df$ADES_PREFIX,1,1) == 'Z' & substring(flights_df$ADES_PREFIX,1,2) != 'ZK' & substring(flights_df$ADES_PREFIX,1,2) != 'ZM', 'Z', substring(flights_df$ADES_PREFIX,1,2))



#fix flights to/from Russia
flights_df$ADEP_PREFIX <- ifelse(substring(flights_df$ADEP_PREFIX,1,1) == 'U' | substring(flights_df$ADEP_PREFIX,1,2) == 'UA' | substring(flights_df$ADEP_PREFIX,1,2) == 'UB' |substring(flights_df$ADEP_PREFIX,1,2) == 'UC' | substring(flights_df$ADEP_PREFIX,1,2) == 'UD'| substring(flights_df$ADEP_PREFIX,1,2) == 'UG' | substring(flights_df$ADEP_PREFIX,1,2) == 'UK' |substring(flights_df$ADEP_PREFIX,1,2) == 'UM' | substring(flights_df$ADEP_PREFIX,1,2) == 'UT' , 'U', substring(flights_df$ADEP_PREFIX,1,2))

flights_df$ADES_PREFIX <- ifelse(substring(flights_df$ADES_PREFIX,1,1) == 'U' | substring(flights_df$ADES_PREFIX,1,2) == 'UA' | substring(flights_df$ADES_PREFIX,1,2) == 'UB' |substring(flights_df$ADES_PREFIX,1,2) == 'UC' | substring(flights_df$ADES_PREFIX,1,2) == 'UD'| substring(flights_df$ADES_PREFIX,1,2) == 'UG' | substring(flights_df$ADES_PREFIX,1,2) == 'UK' |substring(flights_df$ADES_PREFIX,1,2) == 'UM' | substring(flights_df$ADES_PREFIX,1,2) == 'UT' , 'U', substring(flights_df$ADES_PREFIX,1,2))

#fixing flights to/from outermost
flights_df$ADEP_PREFIX<- ifelse( substring(flights_df$ADEP,1,4) == 'FMEE' | substring(flights_df$ADEP,1,4) =='FMEP' | substring(flights_df$ADEP,1,4) =='FMCZ'| substring(flights_df$ADEP,1,4) =='LPCR'|substring(flights_df$ADEP,1,4) == 'LPFL'|substring(flights_df$ADEP,1,4) == 'LPGR'|substring(flights_df$ADEP,1,4) == 'LPHR'|substring(flights_df$ADEP,1,4) == 'LPPD'| substring(flights_df$ADEP,1,4) =='LPLA'|substring(flights_df$ADEP,1,4) == 'LPPI'|substring(flights_df$ADEP,1,4) == 'LPAZ'| substring(flights_df$ADEP,1,4) =='LPSJ'|substring(flights_df$ADEP,1,4) == 'LPMA'|substring(flights_df$ADEP,1,4) == 'LPPS', substring(flights_df$ADEP,1,4), substring(flights_df$ADEP_PREFIX,1,2))

flights_df$ADES_PREFIX<- ifelse( substring(flights_df$ADES,1,4) == 'FMEE' |substring(flights_df$ADES,1,4) == 'FMEP' |substring(flights_df$ADES,1,4) == 'FMCZ'|substring(flights_df$ADES,1,4) == 'LPCR'|substring(flights_df$ADES,1,4) == 'LPFL'|substring(flights_df$ADES,1,4) == 'LPGR'| substring(flights_df$ADES,1,4) =='LPHR'|substring(flights_df$ADES,1,4) == 'LPPD'|substring(flights_df$ADES,1,4) == 'LPLA'| substring(flights_df$ADES,1,4) =='LPPI'|substring(flights_df$ADES,1,4) == 'LPAZ'|substring(flights_df$ADES,1,4) == 'LPSJ'| substring(flights_df$ADES,1,4) =='LPMA'|substring(flights_df$ADES,1,4) == 'LPPS', substring(flights_df$ADEP,1,4), substring(flights_df$ADES_PREFIX,1,2))


colnames(icaoPrefixCategories) <- c("ADEP_PREFIX" ,"ADEP_COUNTRY","ADEP_EU_EEA_EFTA","ADEP_ECAC", "ADEP_OUTERMOST_REGIONS", "ADEP_EU_EEA_EFTA_UK",  "ADEP_OUTER_CLOSE")

str(icaoPrefixCategories)

flights_df <- full_join(flights_df, icaoPrefixCategories)

colnames(icaoPrefixCategories) <- c("ADES_PREFIX" ,"ADES_COUNTRY","ADES_EU_EEA_EFTA","ADES_ECAC", "ADES_OUTERMOST_REGIONS", "ADES_EU_EEA_EFTA_UK",  "ADES_OUTER_CLOSE")

flights_df <- full_join(flights_df, icaoPrefixCategories)

seq <- c('ADEP_PREFIX', 'ADES_PREFIX', 'ADEP_COUNTRY', 'ADEP_EU_EEA_EFTA', 'ADEP_ECAC', 'ADEP_OUTERMOST_REGIONS', 'ADEP_EU_EEA_EFTA_UK','ADEP_OUTER_CLOSE', 'ADES_COUNTRY', 'ADES_EU_EEA_EFTA', 'ADES_ECAC', 'ADES_OUTERMOST_REGIONS', 'ADES_EU_EEA_EFTA_UK', 'ADES_OUTER_CLOSE')
for(colName in seq){
  flights_df[[colName]] <- as.factor(flights_df[[colName]])
  }
return(flights_df)

}
```

```{r}
flights_df <- createCategories(flights_df)
```

```{r}
head(flights_df)
```



### Calculations from fuel tools

```{r}
# Current Exchange rate from Euros to USD
EurosToUsdExchangeRate = 1.182
```


```{r}
CalculateSAFCost <- function(flights_df, costOfSafFuelPerKg = 3.66, safBlendingMandate = 0.02){
  flights_df$SAF_COST <- flights_df$FUEL * safBlendingMandate * costOfSafFuelPerKg
return(flights_df)
}
```

```{r}
flights_df<-CalculateSAFCost(flights_df)
```

```{r}
head(flights_df)
```



```{r}
CalculateFuelCost <- function(flights_df, costOfJetFuelPerKg = 0.61, safBlendingMandate = 0.02 ){
    flights_df$FUEL_COST = flights_df$FUEL * (1-safBlendingMandate) * costOfJetFuelPerKg
    return(flights_df)
}
```

```{r}
flights_df<-CalculateFuelCost(flights_df)
```


```{r}
CalculateTotalFuelCost <- function(flights_df){
    flights_df$TOTAL_FUEL_COST = flights_df$SAF_COST + flights_df$FUEL_COST
    return(flights_df)
    
}
```

```{r}
flights_df<-CalculateTotalFuelCost(flights_df)
```


```{r}
CalculateTaxCost <- function(flights_df, FuelTaxRateEurosPerGJ = 2.15 , blendingMandate=0.02 ){
    # *************************************************** #
    # Constants for Fuel TAX Calculations
    # all prices are in Euros/GJ
    # 2023 = 0 Tax rate
    # 2024 = 1.075    2025 = 2.15 etc

    # Tax rate in 2033
    MaxFuelTaxRateEurosPerGJ <- 10.75
    # Using rate for 2025 to match the SAF mandate





    # Tax rate in Euros/kg
    FuelTaxRateEurosPerKg <- (46.4 / 1000) * FuelTaxRateEurosPerGJ
    FuelTaxRateUsdPerKg <- FuelTaxRateEurosPerKg * EurosToUsdExchangeRate
    # *************************************************** #

    flights_df$TAX_COST = flights_df$FUEL * (1-blendingMandate) * FuelTaxRateUsdPerKg

    return(flights_df)
}
```

```{r}
flights_df<-CalculateTaxCost(flights_df)
```


```{r}
CalculateETSCost <- function(flights_df, safBlendingMandate=0.02, ETSCostpertonne = 62, ETSpercentage = 50){

    ETSPricePerKg <- ETSCostpertonne/1000 * EurosToUsdExchangeRate

    flights_df$ETS_COST <- flights_df$FUEL * 3.15 * (1-safBlendingMandate) * ETSPricePerKg * ETSpercentage/100

    return(flights_df)
}
```

```{r}
flights_df<-CalculateETSCost(flights_df)
```


```{r}
head(flights_df)
```


```{r}
getDFRatio <- function(dfMonthsSet){
    summerIATA <- set(4, 5, 6, 7, 8, 9, 10)
    winterIATA <- set(1, 2, 3, 11, 12)

    reSum <- summerIATA - dfMonthsSet
    reWin <- winterIATA - dfMonthsSet

    sumMultiplier <- len(summerIATA) - len(reSum)
    winMultiplier <- len(winterIATA) - len(reWin)

}
```

```{r}
getIATASeasons <- function(setyear){
    startSummer <- datetime(setyear, 3, 1) + relativedelta.relativedelta(day=31, weekday=relativedelta.SU(-1))

    endSummer <- datetime(setyear, 10, 1) + relativedelta.relativedelta(day=31, hours=24,
                                                                    weekday=relativedelta.SA(-1)) + dt.timedelta(days=1)

}
```


```{r}
head(flights_df)
```


```{r}
flights_df<-write.csv(flights_df,"/Users/bhanukiran/Desktop/Fiverr gig Tashayu/flightsv2.csv", row.names = FALSE)
```
































